\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}

### 2022-04-01 lin-vs-cen.tex
Making progress

\documentclass{patmorin}
\listfiles
\usepackage{pat}
\usepackage{paralist}
\usepackage{dsfont}  % for \mathds{A}
\usepackage[utf8x]{inputenc}
\usepackage{skull}
\usepackage{paralist}
\usepackage{graphicx}
\usepackage[noend]{algorithmic}

@@

@@

\DeclareMathOperator{\VE}{VE}

@@

\maketitle

@@

For a graph $G$ and a vertex $v\in V(G)$, $\deg_G(v):=|\{vw\in E(G)\}|$ denotes the degree of $v$ in $G$.  For a any $S\subseteq V(G)$, $\deg_G(S):=\sum_{v\in S}\deg_G(v)$ is the total degree of $S$.

For a vertex $v$ in a graph $G$, the \emph{open neighbourhood} of $v$ is $N_G(v):=\{w:vw\in E(G)\}$ and the \emph{closed neighbourhood} of $v$ is $N_G[v]:=\{v\}\cup N_G(v)$.  For any $S\subseteq V(G)$, $N_G(s):=\bigcup_{v\in S}N_G(v)$ and $N_G[s]:=\bigcup_{v\in S}N_G[v]$.  The subgraph $G[S]$ of $G$ \emph{induced} by $S$ is the graph with vertex set $V(G[S]):=S\cap V(G)$ and edge set $E(G[S]):=\{vw\in E(G):\{v,w\}\subseteq S\}$.

For a graph $G$ and a partition $\mathcal{P}$ of $V(G)$, the \emph{quotient graph} $G/\mathcal{P}$ is the graph with vertex set $\mathcal{P}$ and that contains an edge $vw$ if and only $G$ contains an edge with one endpoint in $v$ and one endpoint in $w$.  When $\mathcal{P}$ consists of disjoint subsets of $V(G)$ that do not necessarily cover $V(G)$ then we use the convention that $G/\mathcal{P}:=G/\mathcal{P'}$ where $\mathcal{P}'$ is the completion of $\mathcal{P}$ obtained by introducing each missing element of $V(G)$ as singleton set.  More precisely $\mathcal{P'}:=\mathcal{P}\cup\{\{v\}:v\in V(G)\setminus(\cup P)\}$.

A \emph{subdivision} of a graph $G$ is any graph $G'$ that can be obtained from $G$ by replacing each edge $vw$ of $G$ with a path $v,s_1,s_2,\ldots,s_\ell,w$ in which each of the (newly introduced) internal vertices $s_1,\ldots,s_\ell$ have degree exactly $2$.  Each vertex in $V(G')\setminus V(G)$ is called a \emph{subdivision vertex}.  A $\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has exactly $\ell$ internal vertices.  For any graph $G$, we use $G^{\onesub}$ to denote the $1$-subdivision of $G$.  Note that there is a natural bijection between the vertex set $V(G^{\onesub})$ and the vertex/edge set $V(G)\cup E(G)$.  A $\le\!\!\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has at most $\ell$ internal vertices.

\todo[inline]{Surely this next lemma is well-known?}
We make use of the following generalization of Hall's Marriage Theorem \cite{hall:on}:
\begin{lem}\label{hall_vees}
  Let $G$ be a bipartite graph with vertex parts $A$ and $B$ and having the property that, for any $A_0\subseteq A$, $|N_G(W)|\ge 2|A_0|$.  Then $G$ contains a subgraph $\Lambda$ in which each vertex of $A$ has degree exactly two and each vertex of $B$ has degree at most one.
\end{lem}

\begin{proof}
  Let $\Lambda$ be a subgraph of $G$ in which each vertex of $A$ has degree at most two, each vertex of $B$ has degree at most one, and that maximizes $\deg_\Lambda(A)$.  If $\deg_{\Lambda}(A) = 2|\Lambda|$ then there is nothing to prove. Assume therefore, for the sake of contradiction,  that there exists $v_0\in A$ with $\deg_\Lambda(v_0) < 2$.

  We say that a path $v_0,\ldots,v_m$ in $G$ is \emph{$\Lambda$-aternating} if $v_iv_{i+1}\in E(\Lambda)$ for all odd $i\in\{0,\ldots,m-1\}$ and $v_iv_{i+1}\not\in E(\Lambda)$ for all even $i\in\{0,\ldots,m\}$.  Observe that any such path has even length so that it ends at a vertex $v_m\in A$.  Otherwise, replacing the edges $\lfloor m/2\rfloor$ edges $\{v_1v_2,v_3v_4,\ldots,v_{m-2}v_{m-1}\}\in E(\Lambda)$ with the $\lceil m/2\rceil$ edges $\{v_0v_1,v_2v_3,\ldots,v_{m-1}v_m\}$ does not change $\deg_\Lambda(v)$ for any $v\in A_0\setminus\{v_0\}$ but does increase  $\deg_\Lambda(v_0)$, contradicting the assumption that $\Lambda$ maximizes $\deg_\Lambda(A)$.

  Let $A_0\subseteq A$ and $B_0\subseteq B$ be the subsets of $A$ and $B$, respectively, that can be reached from $v_0$ by $\Lambda$-alternating paths.
  Observe that, for any $x\in B_0$,  there is exactly one edge $vx$ of $\Lambda$ incident on $x$ and the other endpoint $v$ of this edge is in $A_0$.  Let $a$ be the number of edges of $\Lambda$ having one endpoint $A_0$ and one endpoint in $B_0$, so that $a=\deg_\Lambda(B_0)=|B_0|$.  Let $b$ be the number of edges of $\Lambda$ having one endpoint in $A_0$ and one endpoint in $B\setminus B_0$, so that $\deg_\Lambda(A_0)=a+b = |B_0|+b$.    Since $\deg_{\Lambda}(v) \le 2$ for each $v\in A_0$ and $\deg_\Lambda(v_0)<2$, $\deg_\Lambda(A_0)<2|A_0|$.  This yields the desired contradiction, since $|N_\Lambda(A_0)| \le \deg_\Lambda(A_0) = |B_0|+b = \deg_\Lambda(A_0)< 2|A_0|$.greater than
\end{proof}

\begin{rem}
  The constant $2$ that appears twice in \cref{hall_vees} can be replaced with any positive integer $c$.
\end{rem}

@@

Refer to \cref{grid_figure}.  For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.  For each $i\in\{1,\ldots,a\}$, the \emph{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.

In this work, we need to work with pseudogrids.  An \defin{$a\times b$-pseudogrid} is any graph that can be obtained from $G_{a\times b}$ in the following way:
\begin{compactenum}
  \item Replace each edge $vw$ of $G_{a\times b}$ with a path $\overline{P}_{vw}$ whose endpoints are $v$ and $w$.  (In other words, $\overline{P}_{vw}$ is a path obtained by subdividing the edge $vw$ zero or more times.)
  \item Replace each degree-$4$ vertex $(x,y)$ of $G_{a\times b}$ with a path $P_v$. If $P_v$ has only one vertex then this operation has no effect.  Otherwise, $P_v$ has two endpoints $p$ and $q$ and
  \begin{compactenum}[(Q1)]
    \item $p$ is adjacent to $\{(x-1,y), (x,y-1)\}$ and $q$ is adjacent to $(x+1,y),(x,y+1)$;
    \item $p$ is adjacent to $\{(x,y+1), (x-1,y)\}$ and $q$ is adjacent to $(x,y-1),(x+1,y)$; or
    \item $p$ is adjacent to $\{(x-1,y),(x+1,y)\}$ and $q$ is adjacent to $\{(x,y-1),(x,y+1)\}$.\label{q_iii}
  \end{compactenum}
\end{compactenum}

Let $G$ be an $a\times b$ pseudogrid.  For an edge $vw$ of $G_{a\times b}$, we let $P_{vw}:=\overline{P}_{vw}-\{v,w\}$ denote the (possibly empty) subpath containing the internal vertices of $\overline{P}_{vw}$.  For each vertex $v$ of $G_{a\times b}$ of degree less than $4$ we define $P_{v}$ to be the $1$-vertex path that contains only $v$.  In this way, $\mathcal{P}:=\{V(P_v):v\in V(G_{a\times b})\}\cup\{V(P_{vw}):vw\in E(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths.  We call $\mathcal{P}$ the \defin{grid-partition} of $G$ and we denote it by $(P_\mu:\mu\in V(G_{a\times b})\cup E(G_{a\times b}))$.  This defines a mapping $\delta:V(G)\to V(G_{a\times b})\cup E(G_{a\times b})$ where $u\in V(P_{\delta(u)})$ for each $u\in v(G)$.

@@

Each row $R':=v_1,\ldots,v_a$ of $G_{a\times b}$ corresponds naturally to a path $R$ of $G$. The path $R$ contains $V(\overline{P}_{v_iv_{i+1}})$ for each $i\in\{1,\ldots,a-1\}$.  However, for $i\in\{2,\ldots,a-1\}$ $R$ may or may not contain $V(P_{v_i})$.  In particular, if $P_{v_i}$ was created using (Q\ref{q_iii}) then $R$ does not contain any internal vertices in $P_{v_i}$.
Similarly, a column $C':=v_1,\ldots,v_b$ of $G_{a\times b}$ corresponds to a path that contains $V(\overline{P}_{v_jv_{j+1}})$ for each $j\in\{1,\ldots,b-1\}$. This correspondence allows us to talk about the rows and columns of $G$, which we will do next.

As part of our proof, we use the operation of \defin{deleting} a row (or column) $R$ of $G$, which works as follows:  We remove all edges of $R$ from $G$ and eliminate any isolated vertices.  If this produces vertices of degree $1$ (which happens when $R$ is the first or last row of $G$) then we repeatedly remove degree-$1$ vertices until none remain.  If $G$ is an $a\times b$ pseudogrid and we delete some row $R$, then the resulting graph is an $a\times (b-1)$ pseudogrid.  Similarly, if we delete column $C$ of $G$, then the resulting graph is an $(a-1)\times b$ pseudogrid.

@@

Let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:=(P_\mu:\mu\in\VE(G_{k\times k}))$ and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$.  For any $\mu\in\VE(G_{k\times k})$, we let $\varphi_{\mathcal{P}}(\mu):=\{\phi(v):v\in V(P_\mu)\}$.  For any colour $\alpha\in\{1,\ldots,c\}$, define $\varphi^{-1}(\alpha):=\{v\in V(G):\varphi(v)=\alpha\}$ and define $\varphi_\mathcal{P}^{-2}(\alpha):=\{\mu\in \VE(G_{k\times k}):\alpha\in\varphi_\mathcal{P}(\mu)$.  For any colour set $A\subseteq\{1,\ldots,c\}$ define $\varphi^{-1}(A):=\bigcup_{\alpha\in A}\varphi^{-1}(\alpha)$ and $\varphi_\mathcal{P}^{-2}(A):=\bigcup_{\alpha\in A}\varphi_{\mathcal{P}}^{-2}(\alpha)$.

@@

\begin{lem}
  Let $0 < \epsilon < 1$, let $k\ge 1/\epsilon$ be an integer, let $G$ be a $k\times k$ pseudogrid and let $\varphi:V(G)\to\{1,\ldots,\lfloor\epsilon k\floor\}$ be a vertex colouring of $G$.
  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ that has a grid-partition $\mathcal{P}':=(P'_\mu:\mu\in \VE(G_{k'\times k'})$ such that
  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-2}(A)| \ge |A|/2\epsilon$.
\end{lem}

\begin{proof}
  The proof is by induction on $|\varphi(V(G))|$, the number of colours used by the colouring $\varphi$.  If there exists no $A\subseteq\{1,\ldots,\lfloor\epsilon k\rfloor\}$ with $\varphi_{\mathcal{P}}^{-2}(A) < |A|/2\epsilon$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma. Now assume that such a set $A$ exists.

  We will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le |A|/2\epsilon$.  For each $v:=(i,j)\in V(G_{k\times k})$ with $A\cap\varphi(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.  For each edge $vw\in E(G_{k\times k})$ with $A\cap\phi()\alpha\in\Phi(vw)$, we include the row of $G$ that contains $P_{vw}$ in $R$ or we include the column of $G$ that contains $P_{vw}$ in $C$.  Finally, we add arbitrary rows to $R$ or columns to $C$ to ensure that $|R|=|C|$.  Now we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-|A|/2\epsilon$ and such that $\varphi(V(G_0))\cap A=\emptyset$.  In particular, $|\varphi(V(G_0))|\le |\varphi(V(G))|-|A|$.

  \todo[inline]{Now apply induction......}
\end{proof}

\begin{lem}
  Let $C>1$ be an integer, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(G)$, $|\varphi_{\mathcal{P}'}^{-2}(A)| \ge C\cdot|A|+1$. Then there exists a partial colouring $\phi:\VE(G_{k\times k})\to\{1,\ldots,c,\perp\}$ with the following properties:
  \begin{compactenum}[(i)]
    \item For each $\mu\in\VE(G_{k\times k})$, $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi(\mu)$.
    \item For each $\alpha\in\varphi(V(G))$, $|\{\mu\in\VE(G_{k\times k}):\phi(\mu)=\alpha\}|\ge C$.
  \end{compactenum}
\end{lem}

\begin{proof}
  Use Hall's Theorem to find $C$ matchings such that each colour $\alpha$ appears in all $C$ matchings and each element of $\VE(G_{k\times k})$ appears in at most one matching.
\end{proof}

\subsection{Finding Paths Through Well-Separated Vertex/Edge Pairs}

The \defin{$r$-box} centered at a vertex $v:=(i,j)$ of $G_{a\times b}$ is defined as
\[  B_r(v) := \{i-r,\ldots,i+r\}\times\{j-r,\ldots,j+r\} \cap V(G_{a,b}) \enspace . \]
The \defin{$r$-box} centered at an edge $vw$ of $G_{a\times b}$ is $B_r(vw):=B_r(v)\cup B_r(w)$.  Observe that, for any $\mu\in V(G_{a\times b})\cup E(G_{a\times b})$,  $v\in B_r(u)$ if and only if $u\in B_r(v)$.  Any $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{a\times b}[B_r(\mu)]$.  We extend these definitions to vertices of an $a\times b$ pseudogrid $G$ so that
\[
   \tilde{B}_r(v) := \bigcup_{\mu\in V(G_r(\delta(v)) \cup E(G_r(\delta(v))} V(P_\mu)
\]
and $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.

>>>>

\documentclass{patmorin}
\listfiles
\usepackage{pat}
\usepackage{paralist}
\usepackage{dsfont}  % for \mathds{A}
\usepackage[utf8x]{inputenc}
\usepackage{skull}
\usepackage{paralist}
\usepackage{graphicx}
\usepackage[noend]{algorithmic}
\usepackage{bbm}  % needed for \mathbbm{1}

@@

\DeclareMathOperator{\binomial}{binomial}

@@

\DeclareMathOperator{\VE}{\mathit{VE}}

@@

\maketitle
\renewcommand{\E}{\mathbb{E}}
\renewcommand{\Pr}{\mathbb{P}}

@@

% For a graph $G$ and a vertex $v\in V(G)$, $\deg_G(v):=|\{vw\in E(G)\}|$ denotes the degree of $v$ in $G$.  For a any $S\subseteq V(G)$, $\deg_G(S):=\sum_{v\in S}\deg_G(v)$ is the total degree of $S$.

% For a vertex $v$ in a graph $G$, the \emph{open neighbourhood} of $v$ is $N_G(v):=\{w:vw\in E(G)\}$ and the \emph{closed neighbourhood} of $v$ is $N_G[v]:=\{v\}\cup N_G(v)$.  For any $S\subseteq V(G)$, $N_G(s):=\bigcup_{v\in S}N_G(v)$ and $N_G[s]:=\bigcup_{v\in S}N_G[v]$.  The subgraph $G[S]$ of $G$ \emph{induced} by $S$ is the graph with vertex set $V(G[S]):=S\cap V(G)$ and edge set $E(G[S]):=\{vw\in E(G):\{v,w\}\subseteq S\}$.

In this paper, all graphs are simple and undirected. For a graph $G$, $V(G)$ denotes the vertex set of $G$, $E(G)$ denotes the edge set of $G$ and $\VE(G):=V(G)\cup E(G)$.

% For a graph $G$ and a partition $\mathcal{P}$ of $V(G)$, the \emph{quotient graph} $G/\mathcal{P}$ is the graph with vertex set $\mathcal{P}$ and that contains an edge $vw$ if and only $G$ contains an edge with one endpoint in $v$ and one endpoint in $w$.  When $\mathcal{P}$ consists of disjoint subsets of $V(G)$ that do not necessarily cover $V(G)$ then we use the convention that $G/\mathcal{P}:=G/\mathcal{P'}$ where $\mathcal{P}'$ is the completion of $\mathcal{P}$ obtained by introducing each missing element of $V(G)$ as singleton set.  More precisely $\mathcal{P'}:=\mathcal{P}\cup\{\{v\}:v\in V(G)\setminus(\cup P)\}$.

% A \emph{subdivision} of a graph $G$ is any graph $G'$ that can be obtained from $G$ by replacing each edge $vw$ of $G$ with a path $v,s_1,s_2,\ldots,s_\ell,w$ in which each of the (newly introduced) internal vertices $s_1,\ldots,s_\ell$ have degree exactly $2$.  Each vertex in $V(G')\setminus V(G)$ is called a \emph{subdivision vertex}.  A $\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has exactly $\ell$ internal vertices.  For any graph $G$, we use $G^{\onesub}$ to denote the $1$-subdivision of $G$.  Note that there is a natural bijection between the vertex set $V(G^{\onesub})$ and the vertex/edge set $V(G)\cup E(G)$.  A $\le\!\!\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has at most $\ell$ internal vertices.

We make use of the following (polygamous) generalisation of Hall's Marriage Theorem, which follows from Hall's Marriage Theorem by replacing each vertex $x\in X$ with a $d$-tuple $x_1,\ldots,x_d$ each having the same neighbours as $x$.

% \begin{lem}\label{normal_hall}
%   Let $G$ be a bipartite graph with parts $X$ and $Y$ such that, for any $A\subseteq X$, $|N_G(A)|\ge |A|+1$.  Then $G$ contains a subgraph $M$ (a matching) such that $\deg_M(x)=1$ for each $x\in X$ and $\deg_M(y)\le 1$ for each $y\in Y$.
% \end{lem}

\begin{cor}\label{d_hall}
  Let $d\ge 1$ be an integer and let $G$ be a bipartite graph with parts $X$ and $Y$ such that, for any $A\subseteq X$, $|N_G(A)|\ge d|A|$.  Then $G$ contains a subgraph $M$ such that $\deg_M(x)=d$ for each $x\in X$ and $\deg_M(y)\le 1$ for each $y\in Y$.
\end{cor}

\begin{lem}\label{weighted_lovasz}
  Let $E:=\{E_1,\ldots,E_n\}$ be a set of events in some probability space $(\Omega,\Pr)$.  For each $i\in\{1,\ldots,n\}$, let $\Gamma_i\subseteq E$ be such that the event $E_i$ is mutually independent of $\Gamma_i$,\footnote{An event $A$ is mutually independent of a set $\{B_1,\ldots,B_r\}$ of events if, for any disjoint sets $I,J\subseteq\{1,\ldots,r\}$, $\Pr(A\cap\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j=\Pr(A)\Pr(\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j)$.} and let $w:E\to(0,1)$ be such that
  \[
      \Pr(E_i) \le w(E_i)\cdot\prod_{E_j\in\Gamma_i}(1-w(E_j))  \enspace .
  \]
  Then $\Pr(\overline{E_1}\cap\cdots\cap\overline{E_n}) > 0$.
\end{lem}

% \begin{proof}
%   \todo[inline]{Surely this is known}
% \end{proof}
%
%
% \todo[inline]{Surely this next lemma is well-known?}
% We make use of the following generalization of Hall's Marriage Theorem \cite{hall:on}:
% \begin{lem}\label{hall_vees}
%   Let $G$ be a bipartite graph with vertex parts $A$ and $B$ and having the property that, for any $A_0\subseteq A$, $|N_G(W)|\ge 2|A_0|$.  Then $G$ contains a subgraph $\Lambda$ in which each vertex of $A$ has degree exactly two and each vertex of $B$ has degree at most one.
% \end{lem}
%
% \begin{proof}
%   Let $\Lambda$ be a subgraph of $G$ in which each vertex of $A$ has degree at most two, each vertex of $B$ has degree at most one, and that maximizes $\deg_\Lambda(A)$.  If $\deg_{\Lambda}(A) = 2|\Lambda|$ then there is nothing to prove. Assume therefore, for the sake of contradiction,  that there exists $v_0\in A$ with $\deg_\Lambda(v_0) < 2$.
%
%   We say that a path $v_0,\ldots,v_m$ in $G$ is \emph{$\Lambda$-aternating} if $v_iv_{i+1}\in E(\Lambda)$ for all odd $i\in\{0,\ldots,m-1\}$ and $v_iv_{i+1}\not\in E(\Lambda)$ for all even $i\in\{0,\ldots,m\}$.  Observe that any such path has even length so that it ends at a vertex $v_m\in A$.  Otherwise, replacing the edges $\lfloor m/2\rfloor$ edges $\{v_1v_2,v_3v_4,\ldots,v_{m-2}v_{m-1}\}\in E(\Lambda)$ with the $\lceil m/2\rceil$ edges $\{v_0v_1,v_2v_3,\ldots,v_{m-1}v_m\}$ does not change $\deg_\Lambda(v)$ for any $v\in A_0\setminus\{v_0\}$ but does increase  $\deg_\Lambda(v_0)$, contradicting the assumption that $\Lambda$ maximizes $\deg_\Lambda(A)$.
%
%   Let $A_0\subseteq A$ and $B_0\subseteq B$ be the subsets of $A$ and $B$, respectively, that can be reached from $v_0$ by $\Lambda$-alternating paths.
%   Observe that, for any $x\in B_0$,  there is exactly one edge $vx$ of $\Lambda$ incident on $x$ and the other endpoint $v$ of this edge is in $A_0$.  Let $a$ be the number of edges of $\Lambda$ having one endpoint $A_0$ and one endpoint in $B_0$, so that $a=\deg_\Lambda(B_0)=|B_0|$.  Let $b$ be the number of edges of $\Lambda$ having one endpoint in $A_0$ and one endpoint in $B\setminus B_0$, so that $\deg_\Lambda(A_0)=a+b = |B_0|+b$.    Since $\deg_{\Lambda}(v) \le 2$ for each $v\in A_0$ and $\deg_\Lambda(v_0)<2$, $\deg_\Lambda(A_0)<2|A_0|$.  This yields the desired contradiction, since $|N_\Lambda(A_0)| \le \deg_\Lambda(A_0) = |B_0|+b = \deg_\Lambda(A_0)< 2|A_0|$.greater than
% \end{proof}
%
% \begin{rem}
%   The constant $2$ that appears twice in \cref{hall_vees} can be replaced with any positive integer $c$.
% \end{rem}

@@

% Refer to \cref{grid_figure}.
For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.  For each $i\in\{1,\ldots,a\}$, the \emph{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.

In this work, we need to work with pseudogrids.  An \defin{$a\times b$-pseudogrid} is any graph that can be obtained from $G_{a\times b}$ in the following way:
\begin{compactenum}
  \item Replace each edge $vw$ of $G_{a\times b}$ with a path $\overline{P}_{vw}$ whose endpoints are $v$ and $w$.  (In other words, $\overline{P}_{vw}$ is a path obtained by subdividing the edge $vw$ zero or more times.)
  \item Replace each degree-$4$ vertex $(x,y)$ of $G_{a\times b}$ with a path $P_v$. If $P_v$ has only one vertex then this operation has no effect.  Otherwise, $P_v$ has two endpoints $p$ and $q$ and
  \begin{compactenum}[(Q1)]
    \item \label{q_i} $p$ is adjacent to $\{(x-1,y), (x,y-1)\}$ and $q$ is adjacent to $(x+1,y),(x,y+1)$;
    \item \label{q_ii} $p$ is adjacent to $\{(x,y+1), (x-1,y)\}$ and $q$ is adjacent to $(x,y-1),(x+1,y)$; or
    \item \label{q_iii} $p$ is adjacent to $\{(x-1,y),(x+1,y)\}$ and $q$ is adjacent to $\{(x,y-1),(x,y+1)\}$.
  \end{compactenum}
\end{compactenum}

Let $G$ be an $a\times b$ pseudogrid.  For an edge $vw$ of $G_{a\times b}$, we let $P_{vw}:=\overline{P}_{vw}-\{v,w\}$ denote the (possibly empty) subpath containing the internal vertices of $\overline{P}_{vw}$.  For each vertex $v$ of $G_{a\times b}$ of degree less than $4$ we define $P_{v}$ to be the $1$-vertex path that contains only $v$.  In this way, $\mathcal{P}:=\{V(P_\mu):\mu\in \VE(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths.  We call $\mathcal{P}$ the \defin{grid-partition} of $G$.
% and we denote it by $(P_\mu:\mu\in \VE(G_{a\times b}))$.
% This defines a mapping $\delta:V(G)\to V(G_{a\times b})\cup E(G_{a\times b})$ where $u\in V(P_{\delta(u)})$ for each $u\in v(G)$.

@@

Each row $R':=v_1,\ldots,v_a$ of $G_{a\times b}$ corresponds naturally to a path $R$ of $G$. The path $R$ contains $V(\overline{P}_{v_iv_{i+1}})$ for each $i\in\{1,\ldots,a-1\}$.  However, for $i\in\{2,\ldots,a-1\}$ $R$ may or may not contain $V(P_{v_i})$.  In particular, if $P_{v_i}$ was created using (Q\ref{q_iii})\todo{Fix this!} then $R$ does not contain any internal vertices in $P_{v_i}$.
Similarly, a column $C':=v_1,\ldots,v_b$ of $G_{a\times b}$ corresponds to a path that contains $V(\overline{P}_{v_jv_{j+1}})$ for each $j\in\{1,\ldots,b-1\}$. This correspondence allows us to talk about the rows and columns of $G$, which we will do next.

As part of our proof, we use the operation of \defin{deleting} a row (or column) $R$ of $G$, which works as follows:  We remove all edges of $R$ from $G$ and eliminate any isolated vertices.  If this produces vertices of degree $1$ (which happens when $R$ is the first or last row of $G$ or when $R=v_1,\ldots,v_a$ does not contains $P_{v_i}$ for some $i\in\{1,\ldots,r\}$) then we repeatedly remove degree-$1$ vertices until none remain.  If $G$ is an $a\times b$ pseudogrid and we delete some row $R$, then the resulting graph is an $a\times (b-1)$ pseudogrid.  Similarly, if we delete column $C$ of $G$, then the resulting graph is an $(a-1)\times b$ pseudogrid.

@@

For convenience, let $G_k:=G_{k\times k}$.
Let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:=(P_\mu:\mu\in\VE(G_{k}))$ and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$.  For any $\mu\in\VE(G_{k})$, we let $\varphi_{\mathcal{P}}(\mu):=\{\varphi(v):v\in V(P_\mu)\}$.  For any colour $\alpha\in\{1,\ldots,c\}$, define $\varphi^{-1}(\alpha):=\{v\in V(G):\varphi(v)=\alpha\}$ and define $\varphi_\mathcal{P}^{-1}(\alpha):=\{\mu\in \VE(G_{k\times k}):\alpha\in\varphi_\mathcal{P}(\mu)\}$.  For any colour set $A\subseteq\{1,\ldots,c\}$ define $\varphi^{-1}(A):=\bigcup_{\alpha\in A}\varphi^{-1}(\alpha)$ and $\varphi_\mathcal{P}^{-1}(A):=\bigcup_{\alpha\in A}\varphi_{\mathcal{P}}^{-1}(\alpha)$.

@@

\begin{lem}\label{only_frequent}
  Let $d\ge 1$ and $k\ge 1$ be integers, let $G$ be a $k\times k$ pseudogrid and let $\varphi$ be a vertex colouring of $G$ that uses $|\varphi(V(G))|\le k/d$ colours.
  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - d|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)| \ge d|A|$.
\end{lem}

\begin{proof}
  The proof is by induction on $|\varphi(V(G))|$, the number of colours used by the colouring $\varphi$.  If there exists no $A\subseteq \varphi(V(G))$ with $\varphi_{\mathcal{P}}^{-1}(A) < d|A|$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma. Otherwise, we will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le d|A|$ to eliminate all vertices with colours in $A$, as follows:
  \begin{compactitem}
    \item For each $v:=(i,j)\in V(G_{k})$ with $A\cap\varphi(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
    \item For each edge $vw\in E(G_{k\times k})$ with $A\cap\phi()\alpha\in\Phi(vw)$, we include the row of $G$ that contains $P_{vw}$ in $R$ or we include the column of $G$ that contains $P_{vw}$ in $C$.
    \item Finally, we add arbitrary rows to $R$ or columns to $C$ to ensure that $|R|=|C|$.
  \end{compactitem}
  Now we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-d|A|$ and such that $\varphi(V(G_0))\cap A=\emptyset$.  In particular, $|\varphi(V(G_0))|\le |\varphi(V(G))|-|A|$.

  Now apply induction on $G_0$ to get a $k'\times k'$ pseudogrid with
  \[  k'\ge k_0-d|\varphi(V(G_0))|  \ge k-d|A|-d|\varphi(V(G_0))|= k - d|\varphi(V(G))|
  \]
  that satisfies the conditions of the lemma.
\end{proof}

\begin{lem}\label{one_colour_per_object}
  Let $d>1$ be an integer, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}'}^{-1}(A)| \ge d\cdot|A|$. Then there exists a (partial) colouring $\phi:\VE(G_{k})\to\{1,\ldots,c,\perp\}$ with the following properties:
  \begin{compactenum}[(i)]
    \item For each $\mu\in\VE(G_{k})$, $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi(\mu)$.
    \item For each $\alpha\in\varphi(V(G))$, $|\{\mu\in\VE(G_{k}):\phi(\mu)=\alpha\}|\ge d$.
  \end{compactenum}
\end{lem}

\begin{proof}
  Consider the bipartite graph $H$ with parts $X:=\{1,\ldots,c\}$ and $Y:=\VE(G_k)$ and edge set
  \[
    \{ (\alpha,\mu)\in X\times Y: \alpha\in\varphi(\mu) \}
  \]
  By \cref{d_hall}, $H$ contains a subgraph $M$ with $\deg_M(x)=d$ for each $\alpha\in X$ and $\deg_M(y)\le 1$ for each $\mu\in Y$.  For each edge $\alpha\mu$ in $M$, set $\phi(\mu):=\alpha$.  This defines $\phi(\mu)$ for any $\mu\in X$ with $\deg_M(\mu)=1$.  For each $\mu\in X$ with $\deg_M(\mu)=0$ set $\phi(\mu):=\perp$.
\end{proof}

\subsection{Finding Paths Through Well-Separated Pairs}

Next we show that, given a sufficiently ``well-separated'' set $S$ of pairs of vertices in $G$, we can always find a path in $G$ that contains every vertex in $S$. For this, we need some definitions of boxes in $G$ and in $G_k$.
